---
title: "p8105_hw1_yl5830"
author: "Yunjia Liu"
date: "2024-09-21"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(ggplot2)
```

## Question1

### 1.Load the penguins dataset
```{r}
data("penguins", package = "palmerpenguins")
penguins_df <- penguins
```

### 2.Describe the penguins dataset
describe the dataset from several aspects, including the head of the dataframe using head() function, the size of the dataframe using nrow() & ncol().
In addition,calculate the mean of flipper_length_mm using mean() function.(Note: already exclude all the possible missing numbers in thr column of flipper_length_mm )

```{r}
head(penguins,5)

num_rows = nrow(penguins_df)
num_cols = ncol(penguins_df)
cat("The dataset has", num_rows, "rows and", num_cols, "columns.\n")


mean_flipper_length = mean(penguins_df$flipper_length_mm, na.rm = TRUE)
cat("The mean flipper length is", mean_flipper_length, "mm.\n")
```

### 3. Scatterplot of flipper_length_mm (y) vs bill_length_mm (x) with color points using the species variable

3.1 Remove rows with missing values to avoid errors in plotting
```{r}
penguins_df_clean = na.omit(penguins_df)
```

3.2 Draw the scatterplot of flipper_length_mm vs bill_length_mm
```{r}
flipper_length_mm, color = species)) +
  geom_point() +
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Flipper Length vs Bill Length by Species") +
  theme_minimal()

print(scatterplot_flipper_bill)
```

3.3 save the plot to the working directory
```{r}
ggsave(filename = "flipper_vs_bill_length_plot.png", plot = scatterplot_flipper_bill, width = 8, height = 6)
```

